---
title: "CCBR-1472: Franchini SAMT derivatives (Arif, Emmanuel) - For MS"
author: "Kate Goldfarbmuren"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: FALSE
editor_options:
  chunk_output_type: console
params:
  
  
  #   ###### for biowulf downtime
  
  
  study_name: "ccbr1472_SAMTderivatives_KCG_ForMS"
    ##### where is the data
    #### usual place on biowulf
  parent_dir: "~/../../Volumes/NCI_VB/franchini/ccbr1472_SAMTderivatives_Arif_KCG/ForMS/"
  input_dir: "~/../../Volumes/NCI_VB/franchini/ccbr1472_SAMTderivatives_Arif_KCG/ForMS/input_data"
    #### where is the code
  code_path: "~/../../Volumes/NCI_VB/goldfarbmurenkc/github/VaccineBranch/Franchini/ccbr1472_SAMTderivatives_Arif_ForMS_KCG.Rmd" ### this file
  pval_cutoff: "0.05" 
  lib_loc_to_use: "/data/goldfarbmurenkc/R/rhel8/4.5/"
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.width = 7, 
                      fig.height = 7, 
                      dev = "png", #png or cairo_pdf
                      #dpi = 400,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE)

#https://github.com/rstudio/DT/issues/867
DT::datatable(
  matrix(), extensions="Buttons",
  options = list(dom = 'Blfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                 iDisplayLength = 25,
                 lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All")))
)
     
```


```{r generate_dirs, warning=FALSE, include=FALSE}
############################################################
# generate dirs
############################################################
parent_dir = params$parent_dir
input_dir = params$input_dir

outputIMAGES_dir = file.path(parent_dir,"IMAGES")
outputTABLES_dir = file.path(parent_dir,"TABLES")
outputRData_dir = file.path(parent_dir,"RData")
outputEnrichr_dir = file.path(parent_dir,"Enrichr")
output_dirs <- c(outputIMAGES_dir, outputTABLES_dir, outputRData_dir, outputEnrichr_dir)

for (curr_dir in output_dirs){
  if(!(file.exists(file.path(curr_dir)))){
    dir.create(file.path(curr_dir), showWarnings = TRUE)
  }
}

```

```{r source_pkgs, warning=FALSE, include=FALSE}
############################################################
# packages & functions
############################################################
# package list

list.of.packages=c("openxlsx","readxl", "tidyverse","cowplot", "pheatmap","dendsort","ggpubr","stringr","eulerr", "enrichR","ggsurvfit","OlinkAnalyze","knitr","broom","kableExtra","gtsummary",
"gridExtra","DT","downloadthis","ggbeeswarm","ggrepel","plotly","ggplotify","flextable","PCAtools","officer",
"DESeq2","RColorBrewer","colorspace","corrplot","factoextra","data.table","strex","pander",#"plotly",
"ggalluvial","geepack","janitor","tidyxl",
"igraph","networkD3","circlize","psych","scales","rstatix","htmltools")

#install as needed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

# # source common functions, need to have NCI_VB mounted to run this
# source(file.path(params$functions_path,"CommonFunctions_KCG.R"))
# 
# # Rmd functions
# source(file.path(params$functions_path,'Rmd_functions_KCG.R'))

# load packages
invisible(lapply(list.of.packages, library, character.only = TRUE))
```



```{r loadData}

    xlsx_file_path <- file.path(input_dir,paste0(str_before_last(params$study_name,"_"),"_list_of_RAWdatasets.xlsx"))
tab_names<-openxlsx::getSheetNames(xlsx_file_path)
### list with separate element for each tab
rawdatasets_list <- lapply(tab_names,
                           openxlsx::read.xlsx,xlsxFile=xlsx_file_path)
names(rawdatasets_list) <- tab_names




########### Study 3 processing
rawdata_df <- rawdatasets_list[["Study-3"]] %>%

  ### make longer
  tidyr::pivot_longer(!c(treatment,AnimalID), names_to = "cat_var", values_to = "amount", values_transform = as.numeric) %>%
  ### add useful metadata
  dplyr::mutate(
                category = str_before_first(cat_var,"__"),
                variable = str_after_first(cat_var,"__")
  ) 


########## now remove UNSTIM, and take delta vs GP120
phen_long <- rawdata_df %>%
  dplyr::filter(treatment != "UNSTIM") %>%
  dplyr::arrange(variable, AnimalID, treatment) %>%
  dplyr::group_by(variable, AnimalID) %>%
  dplyr::mutate(delta = amount - amount [1L]) %>%
  dplyr::ungroup() %>%
  ### remove gp120 since all zeros
  dplyr::filter(treatment != "GP120") 

#### also make phen_long for gp120 vs unstim
phen_long_stim <- rawdata_df %>%
  dplyr::filter(treatment %in% c("UNSTIM","GP120")) %>%
  dplyr::arrange(variable, AnimalID, desc(treatment)) %>%
  dplyr::group_by(variable, AnimalID) %>%
  dplyr::mutate(delta = amount - amount [1L]) %>%
  dplyr::ungroup() %>%
  ### remove UNSTIM since all zeros
  dplyr::filter(treatment != "UNSTIM") 



list_of_stats <- list()
list_of_plotted <- list()

```

# {.tabset}

## Figure 1 {.tabset}

### Fig1c {.tabset}

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across all pairwise group contrasts*


```{r Fig1c_prep}
group_var_to_use <- "Derivative"
group_colors_to_use <- c("SAMT247"="darkgreen",
                         "MetA"="deepskyblue",
                         "MetB"="black",
                         "MetC"="blue",
                         "MetD"="orange"
                         )
names_for_gtc <- "Fig1c"
cat_to_use_name <- names_for_gtc
curr_ds <- "log10"
skip_ylim_ranges <- TRUE

groups_to_compare <- data.frame("group1"=c("SAMT247","SAMT247","SAMT247","SAMT247",
                                           "MetB","MetC","MetD",
                                           "MetC","MetD",
                                           "MetD"),                            
                                  "group2"=c("MetA","MetB","MetC","MetD",
                                            "MetA","MetA","MetA",
                                            "MetB","MetB",
                                            "MetC")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- rawdatasets_list[["Study-1"]] %>%
    ### only VagSecretion data
    dplyr::select(AnimalID, ends_with("VagSecretion")) %>%
    ##make longer
    tidyr::pivot_longer(!AnimalID, names_to = "deri_var", values_to = "amount") %>%
        ### remove rows with  NAamount (these were since only one rep for VagSecretion measurements)
    dplyr::filter(!is.na(amount)) %>%
    ### separate Derivative and variable
    dplyr::mutate(Derivative = str_before_first(deri_var,"__"),
                  variable = str_after_first(deri_var,"__in"))
    
  
  
vars_to_use <- unique(curr_phen_long$variable)
linewidth_to_use <- 0 ### zero = no lines, non zero = yes add lines



### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
row_font_size <- 14



annotation_row <- data.frame("rowname"=vars_to_use,
                             "dir"=rep("yes",length(vars_to_use))) %>%
  column_to_rownames()


#tmp 2 <- columns = animals, rows = variables
tmp2 <- curr_phen_long %>%

  ### combine animal_der
  dplyr::mutate(animal_der = str_remove_all(paste0(AnimalID,"__",Derivative),"[^\\S ]"),
                ##  log10
                log10_amount = log10(amount)) %>%
    ## only columns need
  dplyr::select(variable,animal_der, log10_amount) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "animal_der", values_from = "log10_amount") %>%
  column_to_rownames("variable")


shape_values_to_use <-c(21,22,23,24,25,1)

```

```{r scatterplots, results='asis', fig.height=6, fig.width=10}
    
  #################### add scatter plots
    
    # cat("#### ", paste0("scatterplots {.active}"),"\n")
    # cat("\n")
    # 
    #### for each sig different variable (or including not sig)
    if(plot_not_sig == "yes"){
      vars_to_plot <- rownames(annotation_row) 
    }else{
      vars_to_plot <- rownames(annotation_row)[which(annotation_row[,1] != "not_sig")]   
    }
    
    ### -------- y-axis range control --------
    ### defaults to applying the shared candidate ranges; specific _prep chunks
    ### (Fig1c/Fig1d/Fig1f) set skip_ylim_ranges <- TRUE beforehand to opt out.
    ### `exists()` check + immediate rm() means this never leaks into later chunks
    ### that don't set it explicitly.
    if (exists("skip_ylim_ranges") && isTRUE(skip_ylim_ranges)) {
      apply_ylim_ranges <- FALSE
    } else {
      apply_ylim_ranges <- TRUE
    }

### number of candidate buckets to cluster into - defaults to 3, but a _prep
    ### chunk (e.g. setVarsScatter_SuppFig3BCD_Mono) can override by setting
    ### n_ylim_buckets beforehand
    if (exists("n_ylim_buckets")) {
      n_buckets_to_use <- n_ylim_buckets
    } else {
      n_buckets_to_use <- 3
    }


##### function for the breaks
get_zero_anchored_breaks <- function(lo, hi, min_ticks = 4, target_ticks = 5) {
  ### lo should be <= 0 <= hi; if not (e.g. all-positive or all-negative data), handle separately
  if (lo > 0) lo <- 0  ### force 0 into range if data happens to be all positive
  if (hi < 0) hi <- 0  ### force 0 into range if data happens to be all negative
  
  full_span <- hi - lo
  
  ### use pretty() to find a nice step size based on the full span, but anchor at 0
  ### pretty() on c(lo, hi) gives evenly spaced nice numbers - we just need to make sure
  ### they include 0 and satisfy min_ticks
  raw_breaks <- pretty(c(lo, hi), n = target_ticks)
  step <- diff(raw_breaks)[1]
  
  ### generate ticks as multiples of step, covering at least [lo, hi]
  n_below <- ceiling(abs(lo) / step)
  n_above <- ceiling(hi / step)
  ticks <- seq(-n_below * step, n_above * step, by = step)
  
  ### if pretty() gave us fewer than min_ticks, shrink the step and retry
  while (length(ticks) < min_ticks) {
    step <- step / 2
    n_below <- ceiling(abs(lo) / step)
    n_above <- ceiling(hi / step)
    ticks <- seq(-n_below * step, n_above * step, by = step)
  }
  
  ticks
}
    ### clear the opt-out flags immediately so they don't silently persist
    if (exists("skip_ylim_ranges")) rm(skip_ylim_ranges)
    if (exists("n_ylim_buckets")) rm(n_ylim_buckets)
    if (exists("manual_ylim_ranges") && !apply_ylim_ranges) rm(manual_ylim_ranges)
    
    
if (apply_ylim_ranges) {
      
      ### if a _prep chunk has set manual_ylim_ranges, use those directly and skip
      ### all the automatic min/max/clustering logic below entirely
      if (exists("manual_ylim_ranges")) {
        candidate_ranges <- manual_ylim_ranges
        rm(manual_ylim_ranges)
        
      } else {
        ### compute per-variable min, max, and abs-spread (for clustering)
        var_min <- apply(tmp2[vars_to_plot, , drop = FALSE], 1, function(x) min(x, na.rm = TRUE))
        var_max <- apply(tmp2[vars_to_plot, , drop = FALSE], 1, function(x) max(x, na.rm = TRUE))
        var_abs_spread <- pmax(abs(var_min), abs(var_max))
        
        ### drop any -Inf/NaN from all-NA rows, and zero-spread rows (log10(0) issue)
        keep <- is.finite(var_min) & is.finite(var_max) & is.finite(var_abs_spread) & var_abs_spread > 0
        var_min <- var_min[keep]
        var_max <- var_max[keep]
        var_abs_spread <- var_abs_spread[keep]
        
        nice_round_signed <- function(x) {
          s <- sign(x)
          a <- abs(x)
          out <- if (a <= 1) ceiling(a * 10) / 10
                 else if (a <= 10) ceiling(a)
                 else if (a <= 100) ceiling(a / 5) * 5
                 else ceiling(a / 10) * 10
          s * out
        }
        
        if (length(var_abs_spread) == 0) {
          candidate_ranges <- list(c(-10, 10))
          
        } else if (length(var_abs_spread) < n_buckets_to_use || length(unique(var_abs_spread)) < n_buckets_to_use) {
          ord <- order(var_abs_spread)
          candidate_ranges <- lapply(ord, function(i) {
            c(nice_round_signed(var_min[i] * 1.1), nice_round_signed(var_max[i] * 1.1))
          })
          
        } else {
          n_clusters <- min(n_buckets_to_use, length(unique(var_abs_spread)))
          log_spreads <- log10(var_abs_spread)
          set.seed(123)
          clust <- kmeans(log_spreads, centers = n_clusters)
          
          cluster_min <- tapply(var_min, clust$cluster, min)
          cluster_max <- tapply(var_max, clust$cluster, max)
          cluster_spread_order <- order(tapply(var_abs_spread, clust$cluster, max))
          
          candidate_ranges <- lapply(cluster_spread_order, function(i) {
            c(nice_round_signed(cluster_min[i] * 1.1), nice_round_signed(cluster_max[i] * 1.1))
          })
        }
      }
      
      print(candidate_ranges)
    }
    ### -------- end y-axis range control --------


      #### list of plots with each variable separately
      list_of_plots <- list()
      plotted_data_df <- data.frame()
      stats_data_df <- data.frame()



      for(curr_var in vars_to_plot){
        
        
        desired_order <- names(group_colors_to_use)
        
        df_to_plot <- tmp2 %>%
          rownames_to_column() %>%
          dplyr::filter(rowname == curr_var ) %>%
          ### make longer
          tidyr::pivot_longer(cols = !rowname,
                              names_to = "AnimalID_Group",
                              values_to = "value") %>%
          ### factorize
          dplyr::mutate(treatment = factor(str_after_first(AnimalID_Group,"__"),desired_order),
                        AnimalID = str_before_first(AnimalID_Group,"__"),
                        AnimalID_Group = NULL)
        
        
        if (apply_ylim_ranges) {
          ### pick the smallest candidate range that fully contains this variable's data
          data_min <- min(df_to_plot$value, na.rm = TRUE)
          data_max <- max(df_to_plot$value, na.rm = TRUE)
          
          range_to_use <- candidate_ranges[[length(candidate_ranges)]]
          for (r in candidate_ranges) {
            if (data_min >= r[1] && data_max <= r[2]) {
              range_to_use <- r
              break
            }
          }
          
          ### always show at least 4 evenly-spaced ticks, anchored at 0, across the chosen range
          breaks_to_use <- get_zero_anchored_breaks(range_to_use[1], range_to_use[2])
          
          ### extend range to cover the ticks - this defines the FIXED, shared visible axis
          ### for every variable that lands in this same bucket
          range_to_use <- c(min(range_to_use[1], min(breaks_to_use)),
                             max(range_to_use[2], max(breaks_to_use)))
          
          axis_span <- diff(range_to_use)
          
          ### reserve a FIXED fraction of the tick-covering range as headroom for brackets.
          ### this headroom is identical for every variable sharing this bucket, so the
          ### visible ylim (ylim_for_plot) NEVER varies across panels sharing range_to_use -
          ### brackets scale to fit inside this band rather than growing the axis to fit them.
          bracket_headroom_frac <- 0.18   ### tune: fraction of axis_span reserved above top tick
          ylim_for_plot <- c(range_to_use[1], range_to_use[2] + axis_span * bracket_headroom_frac)
        }

        ## modify pval to be format needed by stat_pvalue_manual

#pair_grid <- as.data.frame(t(combn(desired_order, 2)), stringsAsFactors = FALSE)

## restrict to the same comparisons used in diffBetweenGroups not all pairwise combos
 pair_grid <- groups_to_compare %>%
   dplyr::select(group1, group2)

names(pair_grid) <- c("t1", "t2")

stat.test <- df_to_plot %>%
  pivot_wider(id_cols = AnimalID, names_from = treatment, values_from = value) %>%
  group_modify(~{
    wide <- .x
    out <- lapply(seq_len(nrow(pair_grid)), function(i){
      t1 <- pair_grid$t1[i]; t2 <- pair_grid$t2[i]
      x <- wide[[t1]]; y <- wide[[t2]]
      keep <- complete.cases(x, y)
      p <- if (sum(keep) < 1) NA_real_ else wilcox.test(x[keep], y[keep], paired = TRUE, exact = FALSE)$p.value

      data.frame(
        group1 = t1,
        group2 = t2,
        comparison = paste(t2, "vs", t1),
        Npairs = sum(keep),
        p_raw = p
      )
    })
    bind_rows(out)
  }) %>%

  dplyr::mutate(p_adj = p.adjust(p_raw, method = "BH"),
         p_raw = signif(p_raw,2),
         p_adj = signif(p_adj,2),
         ### be sure to factorize the groups so get correct order
         group1 = factor(group1, levels = desired_order),
         group2 = factor(group2, levels = desired_order)) %>%


# ---- add positions: MUST pass data + formula because your test wasn't made by rstatix ----

  add_xy_position(
    data = df_to_plot,
    formula = value ~ treatment,
    x = "treatment",
    dodge = 0.8   # optional; use if you have dodged geoms
  )  %>%
  #### need to remake the xy since add_xy_pos can't deal witht he factors
  # ensure group1/group2 are strings, then compute correct numeric positions
  mutate(
    group1_chr = as.character(group1),
    group2_chr = as.character(group2),
    xmin_new = match(group1_chr, desired_order),
    xmax_new = match(group2_chr, desired_order)
  ) %>%
  # replace xmin/xmax with the corrected positions (only where matches exist)
  mutate(
    xmin = ifelse(!is.na(xmin_new), xmin_new, xmin),
    xmax = ifelse(!is.na(xmax_new), xmax_new, xmax)
  ) %>%
  # if you want the 'groups' column to reflect the plotted groups, recompute it:
  mutate(groups = paste0(group1_chr, " / ", group2_chr)) %>%
  select(-group1_chr, -group2_chr, -xmin_new, -xmax_new)

stat.test.sig <- stat.test %>%
  filter(!is.na(p_adj) & p_adj <= 0.1) ## show p < 0.1

### -------- axis-based bracket placement (fixed headroom band) --------
### place p-value brackets ENTIRELY within the pre-reserved headroom band above the
### top tick (range_to_use[2] to ylim_for_plot[2]). Brackets scale/compress to fit
### within this fixed band rather than growing the axis - this guarantees every
### variable sharing the same bucket has an IDENTICAL visible y-axis, regardless of
### how many significant comparisons it has.
if (apply_ylim_ranges && nrow(stat.test.sig) > 0) {
  
  headroom_lo <- range_to_use[2]        ### top of the "real" tick-covered range
  headroom_hi <- ylim_for_plot[2]       ### top of the reserved bracket band
  headroom_span <- headroom_hi - headroom_lo
  
  n_brackets <- nrow(stat.test.sig)
  
  stat.test.sig <- stat.test.sig %>%
    dplyr::arrange(xmin, xmax) %>%   ### consistent stacking order
    dplyr::mutate(
      stack_order = dplyr::row_number() - 1,
      ### single bracket sits low in the band; multiple brackets spread evenly across it
      y.position = if (n_brackets == 1) {
        headroom_lo + headroom_span * 0.02
      } else {
        if(n_brackets == 2){
        headroom_lo + headroom_span * (0.05 + 0.5 * stack_order / (n_brackets - 1))
        }else{
          headroom_lo + headroom_span * (0.05 + 0.85 * stack_order / (n_brackets - 1))
        }
      }
    )
}
### -------- end axis-based bracket placement --------
      

        p <- ggplot(df_to_plot, aes(x = treatment, y = value)) +
          geom_hline(
  yintercept = 0,
  linetype = "dashed",
  color = "grey60",
  linewidth = 0.6
)+
          {if (apply_ylim_ranges) coord_cartesian(ylim = ylim_for_plot)} +
          {if (apply_ylim_ranges) scale_y_continuous(breaks = breaks_to_use)} +
          scale_fill_manual(name = paste0("treatment (color)"), values=group_colors_to_use) +
          scale_color_manual(name = paste0("treatment (color)"), values=group_colors_to_use) +
          scale_shape_manual(name = paste0("AnimalID (shape)"), values=shape_values_to_use) +
          geom_boxplot(aes(
            fill=treatment), alpha = 0.6,
            outliers = FALSE ### since adding all points anyway
          )+
          geom_line(
  aes(
    group = AnimalID,
    color = treatment
  ),
  linewidth = linewidth_to_use,
  alpha = 0.4
)+
          geom_beeswarm(aes(
            fill=treatment,
            color = treatment,
            shape = AnimalID
          ),

          cex=cex_for_beeswarm, ### spread in x direction
          corral = "random",
          corral.width = 0.8,
          size = 3.2, stroke = 0.7, alpha = 0.8)+
          ### override shape for fill so can actually fill it
        #  guides(
    guides(shape = guide_legend(override.aes = list(
      fill   = "grey70",   # makes 21–25 visibly filled in the shape legend
      colour = "black",    # keeps outlines consistent
      alpha  = 1,
      size   = 3.2,
      stroke = 0.7
    ))
  ) +
ylab(paste0(curr_var))+
          theme_bw()+

          theme(legend.position="right",
                axis.ticks.x = element_blank(),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = y_axis_size),
                axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                plot.title = element_text(hjust = 0.5), #centers title
                panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                panel.border = element_blank(), axis.line = element_line(colour = "black")
          )+
          #### add p value from sig_assay_df
          #### step.increase set to 0 since y.position above already encodes full stacking
          stat_pvalue_manual(stat.test.sig, label = "p_adj",
  step.increase = 0,
  tip.length = 0.01)
        
        
        

        #   plot(p)

        # plot(p1)
        p1 <- p
        list_of_plots[[curr_var]] <- p1 + theme(legend.position="none")



        cat("\n")
        cat(' \n \n') ### this is the key!

        plotted_data_df <- rbind(plotted_data_df, df_to_plot)
        
        stats_data_df <- rbind(stats_data_df, stat.test %>% 
                                 dplyr::mutate(variable=paste0(curr_var)) %>%
                                 dplyr::relocate(variable))


      }

      ### get the legend
      #https://stackoverflow.com/questions/13649473/add-a-common-legend-for-combined-ggplots
      g_legend<-function(a.gplot){
        tmp <- ggplot_gtable(ggplot_build(a.gplot))
        leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
        legend <- tmp$grobs[[leg]]
        return(legend)}

      ### for empty filler
      #list_of_plots[[length(vars_to_plot)+1]]<-ggplot() + theme_void()
      #list_of_plots[[length(vars_to_plot)+2]]<-g_legend(p1) ### grabs from the last plot
      list_of_plots[[length(vars_to_plot)+1]]<-g_legend(p1) ### grabs from the last plot

      do.call("grid.arrange",c(list_of_plots, ncol=num_columns_for_plot))






      cat("\n")
      cat(' \n \n') ### this is the key!

      list_of_plotted[[paste0(cat_to_use_name,"_diff_",curr_ds)]] <- plotted_data_df


      cat("\n")
      cat(' \n \n') ### this is the key!




    





  ### save pvalues
      
      
        
  list_of_stats[[paste0(cat_to_use_name,"_diff_",curr_ds,names_for_gtc,"p")]] <- stats_data_df  ### scatterplot stats
  
  #### only for non Fig1
  if(!grepl("Fig1",cat_to_use_name)){

 list_of_stats[[paste0(cat_to_use_name,"_diff_",curr_ds,names_for_gtc,"d")]] <- p_value_df1 %>% 
   rownames_to_column("variable")
 
  }




  cat("\n")
  cat(' \n \n') ### this is the key!

```



### Fig1d {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*


```{r Fig1d_prep}
group_var_to_use <- "Derivative"
group_colors_to_use <- c("MetA"="deepskyblue",
                         "MetD"="orange"
                         )
names_for_gtc <- "Fig1d"
cat_to_use_name <- names_for_gtc
curr_ds <- "log10"
skip_ylim_ranges <- TRUE

groups_to_compare <- data.frame("group1"=c("MetD"),                            
                                  "group2"=c("MetA")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- rawdatasets_list[["Study-1"]] %>%
    ### only VagTissue data
    dplyr::select(AnimalID, rep, ends_with("VagTissue")) %>%
    ##make longer
    tidyr::pivot_longer(!c(AnimalID,rep), names_to = "deri_var", values_to = "amount") %>%

    ### separate Derivative and variable
    dplyr::mutate(Derivative = str_before_first(deri_var,"__"),
                  variable = str_after_first(deri_var,"__in"),
                  ### combine rep with var
                  variable = paste0(variable,"__",rep))

vars_to_use <- unique(curr_phen_long$variable)
linewidth_to_use <- 0 ### zero = no lines, non zero = yes add lines



### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
row_font_size <- 14



annotation_row <- data.frame("rowname"=vars_to_use,
                             "dir"=rep("yes",length(vars_to_use))) %>%
  column_to_rownames()


#tmp 2 <- columns = animals, rows = variables
tmp2 <- curr_phen_long %>%

  ### combine animal_der
  dplyr::mutate(animal_der = str_remove_all(paste0(AnimalID,"__",Derivative),"[^\\S ]"),
                ##  log10
                log10_amount = log10(amount)) %>%
    ## only columns need
  dplyr::select(variable,animal_der, log10_amount) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "animal_der", values_from = "log10_amount") %>%
  column_to_rownames("variable")


shape_values_to_use <-c(21,22,23,24,25,1)

```


```{r scatterplots___Fig1d, ref.label=c('scatterplots'), results='asis', fig.height=3.5, fig.width=7}
```







### Fig1f Met-A plasma vs time {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*


```{r Fig1f_prep}
group_var_to_use <- "timepoint"
group_colors_to_use <- c("Baseline"="black",
                         "Week15"="deepskyblue"
                         )
names_for_gtc <- "Fig1f"
cat_to_use_name <- names_for_gtc
curr_ds <- "amount"
skip_ylim_ranges <- TRUE

groups_to_compare <- data.frame("group1"=c("Baseline"),                            
                                  "group2"=c("Week15")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- rawdatasets_list[["Study-2"]] %>%
    ### only MetA in plasma data
    dplyr::select(AnimalID, starts_with("MetA_in_plasma")) %>%
    ##make longer
    tidyr::pivot_longer(!AnimalID, names_to = "var_time", values_to = "amount") %>%

    ### separate timepoint and variable
    dplyr::mutate(variable = str_before_first(var_time,"__"),
                  timepoint = str_after_first(var_time,"__"))

vars_to_use <- unique(curr_phen_long$variable)
linewidth_to_use <- 0 ### zero = no lines, non zero = yes add lines



### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
row_font_size <- 14



annotation_row <- data.frame("rowname"=vars_to_use,
                             "dir"=rep("yes",length(vars_to_use))) %>%
  column_to_rownames()


#tmp 2 <- columns = animals, rows = variables
tmp2 <- curr_phen_long %>%

  ### combine animal_timepoint
  dplyr::mutate(animal_timepoint = str_remove_all(paste0(AnimalID,"__",timepoint),"[^\\S ]")) %>%
    ## only columns need
  dplyr::select(variable,animal_timepoint, amount) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "animal_timepoint", values_from = "amount") %>%
  column_to_rownames("variable")


shape_values_to_use <-c(0,1,2,5,6,8,12,21,22,23,24,25)

```


```{r scatterplots___Fig1f, ref.label=c('scatterplots'), results='asis', fig.height=3, fig.width=5}
```








## Figure 2 {.tabset}


```{r setPhen_amount}

phen_long_to_use <- rawdata_df

### set dataset
ds_to_use <- c("amount") #c("amount","FC","delta")
row_scale_set <- "row" ### rowscaled since not deltas

    
### for delta but not as tiny breakscale
mycols <- colorRampPalette(c("blue4","white","red4"))(1000)
range <- 5
breakscale <- c(-1*range,
                seq(-.9*range,-.0001*range, length.out=length(mycols)/2-1),
                0,
                seq(0.0001*range,.9*range, length.out=length(mycols)/2-1),
                1*range)


        linewidth_to_use <- 0.5 ### non zero = yes add lines
```


```{r setCat__Fig2C_NK}

cat_to_use <- "NK"
cat_to_use_name <- "Fig2C_NK"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes"

```


```{r setGroups___amount}

group_var_to_use <- "treatment"
group_colors_to_use <- c("UNSTIM"="grey70",
                         "GP120"="grey30"
                         )
names_for_gtc <- "Fig2STIM"

groups_to_compare <- data.frame("group1"=c("UNSTIM"),                            
                                  "group2"=c("GP120")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)

#### for heatmap only
phen_long_for_delta <- phen_long_stim


```



### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas

```{r diffBetweenGroups, results='asis', fig.height=8, fig.width=8}

      
 curr_ds <- ds_to_use[1]

  
  ## full_df only has the variables in columns and the factorized group (plus any other covariates if they are in the model)
  
  full_df <- curr_phen_long %>%
    ## trim to only Animal, var and curr_ds
    dplyr::select(c(AnimalID,group_var_to_use,variable,!!rlang::sym(curr_ds))) %>%
    ## make wider
    tidyr::pivot_wider(names_from = "variable", values_from = all_of(curr_ds)) %>%
    ## remove columns with only NA
    dplyr::select(where(function(x) !all(is.na(x)))) %>%
    

    dplyr::rename(Group_to_use=!!rlang::sym(group_var_to_use)) %>%

    ### factorize group
    dplyr::mutate(Group_f = factor(Group_to_use, intersect(
      names(group_colors_to_use),
      unique(data.frame(curr_phen_long)[,group_var_to_use])))) %>%
    dplyr::select(-Group_to_use)
  
  
  
  p_value_df1 <- data.frame("rowname"=colnames(full_df)) %>%
    dplyr::filter(!rowname %in% c("Group_f","AnimalID"))
  
  ######### if have any columns at all
  if(dim(full_df)[2]>2){
    
    
    
    
    for (g in 1:nrow(groups_to_compare)){
      
      
      
      
      ###### only if you have both groups
      if(length(which(full_df$Group_f == groups_to_compare$group1[g])) != 0 &
         length(which(full_df$Group_f == groups_to_compare$group2[g])) != 0){
        
        
        ## wilcoxon for pairwise differences
        
        ## generate matrix with only variables for stats 
        Assay_mat <- as.matrix(full_df[,p_value_df1$rowname])
        
        
        ######## function for paired wilcoxon
        paired_wilcox_p <- function(df, var, group1, group2, id_col = "AnimalID", group_col = "Group_f") {
  d1 <- df[df[[group_col]] == group1, c(id_col, var)]
  d2 <- df[df[[group_col]] == group2, c(id_col, var)]

  # inner join to keep only paired animals
  m <- base::merge(d1, d2, by = id_col, suffixes = c("_g1", "_g2"))

  x <- m[[paste0(var, "_g1")]]
  y <- m[[paste0(var, "_g2")]]

  # drop incomplete pairs for this variable
  keep <- !is.na(x) & !is.na(y)
  x <- x[keep]; y <- y[keep]

  n_pairs <- length(x)
  if (n_pairs < 1) return(list(p = NA_real_, n_pairs = 0))

  p <- stats::wilcox.test(x, y, paired = TRUE, exact = FALSE)$p.value
  list(p = p, n_pairs = n_pairs)
}
        

        vars <- colnames(full_df)[colnames(full_df) %in% p_value_df1$rowname]

out_list <- lapply(vars, function(v) paired_wilcox_p(
  df = full_df,
  var = v,
  group1 = groups_to_compare$group1[g],
  group2 = groups_to_compare$group2[g],
  id_col = "AnimalID",
  group_col = "Group_f"
))

curr_p_raw <- stats::setNames(sapply(out_list, `[[`, "p"), vars)
curr_npairs <- stats::setNames(sapply(out_list, `[[`, "n_pairs"), vars)

### NOTE: BH adjustment is intentionally NOT done here (per-comparison, across variables).
### We instead adjust later, per-variable across the comparisons in groups_to_compare,
### to match the family used for the scatterplot brackets (exploratory/underpowered
### study -> raw p is primary; BH family = the group_pairs tested per variable, not
### the unrelated variables bundled into one figure).


paired_hl <- function(df, var, group1, group2, id_col = "AnimalID", group_col = "Group_f") {
  d1 <- df[df[[group_col]] == group1, c(id_col, var)]
  d2 <- df[df[[group_col]] == group2, c(id_col, var)]
  m <- base::merge(d1, d2, by = id_col, suffixes = c("_g1", "_g2"))

  x <- m[[paste0(var, "_g1")]]
  y <- m[[paste0(var, "_g2")]]

  keep <- !is.na(x) & !is.na(y)
  x <- x[keep]; y <- y[keep]

  if (length(x) < 1) return(NA_real_)

  # effect = (group2 - group1)
  stats::median(y - x, na.rm = TRUE)
}

curr_est <- stats::setNames(sapply(vars, function(v) paired_hl(
  full_df, v,
  groups_to_compare$group1[g],
  groups_to_compare$group2[g],
  id_col = "AnimalID",
  group_col = "Group_f"
)), vars)

## p_adj column is added later (post-loop), so leave a placeholder NA column here
## to preserve downstream structure/order until it's filled in
curr_p_df <- data.frame(
  rowname = vars,
  curr_p_raw = as.numeric(curr_p_raw[vars]),
  curr_p_adj = NA_real_,
  curr_est  = as.numeric(curr_est[vars]),
  curr_npairs = as.integer(curr_npairs[vars])
)

colnames(curr_p_df) <- c(
  "rowname",
  paste0(groups_to_compare$name[g], "__p_raw"),
  paste0(groups_to_compare$name[g], "__p_adj"),
  paste0(groups_to_compare$name[g], "__est"),
  paste0(groups_to_compare$name[g], "__Npairs")
)
       

p_value_df1 <- dplyr::full_join(p_value_df1, curr_p_df, by="rowname") 
        
                }else{
            #### so that these definitely don't get carried over
            curr_p_raw <- NA
            curr_p_adj <- NA
            curr_est <- NA
            curr_x_len <- NA
            curr_y_len <- NA

            #### combine to df

            curr_p_df <- data.frame(matrix(nrow=0, ncol=5))
            colnames(curr_p_df) <- c("rowname",

                                     paste0(groups_to_compare$name[g],"__p_raw"),
                                     paste0(groups_to_compare$name[g],"__p_adj"),
                                     paste0(groups_to_compare$name[g],"__est"),

                                     paste0(groups_to_compare$name[g],"__Npairs"))
            curr_p_df$rowname <- as.character(curr_p_df$rowname)
             ### add these groups to compare to the p_value_df1
      p_value_df1 <- dplyr::full_join(p_value_df1, curr_p_df, by="rowname") 
      

          }

        }


    
      
     
      
      
      
      
  }else{
      
    }
    
    
    
    ### -------- BH adjustment across group_pairs family, per variable --------
    ### this replaces the old per-comparison (across-variables) adjustment.
    ### family = the comparisons listed in groups_to_compare, evaluated separately
    ### for each variable (row) -- matches how brackets are filtered/labeled in
    ### the scatterplots chunk.
    p_raw_cols <- colnames(p_value_df1)[grepl("__p_raw$", colnames(p_value_df1))]
    p_adj_cols <- gsub("__p_raw$", "__p_adj", p_raw_cols)
    
    if (length(p_raw_cols) > 0) {
      for (i in seq_len(nrow(p_value_df1))) {
        raw_vals <- as.numeric(p_value_df1[i, p_raw_cols])
        adj_vals <- stats::p.adjust(raw_vals, method = "BH")
        p_value_df1[i, p_adj_cols] <- adj_vals
      }
    }
    ### -------- end BH adjustment --------
    
    
    ## format p values
    p_value_df1 <- p_value_df1 %>%
      tibble::column_to_rownames() %>%
      ## set sig figs
      dplyr::mutate_all(signif, 2) %>%
      tibble::rownames_to_column()
    
    
    
    
    ## add up/down with sig columns 
    for(curr_comp in groups_to_compare$name){
      curr_df <- p_value_df1[,c("rowname",colnames(p_value_df1)[grepl(paste0(curr_comp,"__"),colnames(p_value_df1))])] %>%
        `colnames<-`(c("rowname","p_raw","p_adj","est","Npairs")) %>%
        ###use p_adj since for single pair of groups will be the same as p_raw
        dplyr::mutate(dir = factor(ifelse(p_adj < 0.05, ifelse(est > 0, "up, p < 0.05","down, p < 0.05"),
                                          ifelse(p_adj < 0.1, ifelse(est > 0, "up, p < 0.1","down, p < 0.1"),
                                                 ifelse(p_adj>0.1,"not_sig","no_test"))),
                                   c("up, p < 0.05","up, p < 0.1","not_sig","down, p < 0.1","down, p < 0.05","no_test"))) %>%
        dplyr::select(-c(p_raw,p_adj,est,Npairs)) %>%
        `colnames<-`(c("rowname",paste0(curr_comp,"__dir")))
      
      p_value_df1 <- dplyr::left_join(p_value_df1, curr_df, by = "rowname")
      
    }
    
    
    ## format p values
    p_value_df1 <- p_value_df1 %>%
      tibble::column_to_rownames() 
    
    
    
    
    
    
    ################################################ NOW plot heatmaps ################################################
     
    sig_assay_df <- p_value_df1 %>%
      tibble::rownames_to_column() %>%
      ### add factored var
      dplyr::mutate(variable = factor(rowname,vars_to_use)) 

          list_of_stats[[paste0(cat_to_use_name,"_diff_",curr_ds,names_for_gtc,"h")]] <- sig_assay_df  ### stats
 
     
   
    #### dir columns
    dir_columns <- rev(colnames(sig_assay_df)[grepl("dir",colnames(sig_assay_df))])
    
    ##### annotation for rows, first listed will be closest to the heatmap (right side of ann cols), last will be on the left edge of the plot
    annotation_row = sig_assay_df[,c(dir_columns,"variable","rowname")] %>%
      ### add simpler rowname with out the annotation
      #dplyr::mutate(rowname_clean = str_before_first(rowname,"xx__")) %>%
      tibble::column_to_rownames() 
    
    
    
    ###### annotation for columns
    annotation_col <- full_df %>%
      dplyr::mutate(animal_group = paste0(AnimalID,"__",as.character(Group_f))) %>%
      tibble::column_to_rownames("animal_group") %>%
      ## arrange
      dplyr::arrange(Group_f,AnimalID) %>%
            dplyr::select(Group_f) 
    

      #### set orders to use
      hm_orders_to_use <- list(
        #"by Assay, dir"=c("variable",dir_columns),
        "by dir, assay"=c(rev(dir_columns),"variable")
      )
      
    
    
       ####
    
    
    annotation_colors=list()
    annotation_colors[["Group_f"]] <- group_colors_to_use

    for(curr_dir in dir_columns){
      annotation_colors[[paste0(curr_dir)]] <- c("up, p < 0.05"="red2","up, p < 0.1"="lightpink",not_sig="black",
                                                 "down, p < 0.1"="skyblue","down, p < 0.05"="blue2",no_test="white")
    }
    
    
    # hmo <- names(hm_orders_to_use)[1]
    
    
      tmp <- full_df %>%
        ### make rownames
        dplyr::mutate(animal_group = paste0(AnimalID,"__",as.character(Group_f))) %>%
      tibble::column_to_rownames("animal_group") %>%
        ##remove Group_f
        dplyr::select(-Group_f,-AnimalID) %>% 
        ## transpose to get columns = animals, rows = variables
        t() %>%
        data.frame() 
    
      #### set up df to collect stats
        plotted_data_df <- data.frame(matrix(ncol=12,nrow=0))
        colnames(plotted_data_df) <- c("rowname",paste0(colnames(tmp)[grepl(names(group_colors_to_use)[2],colnames(tmp))],"vs",names(group_colors_to_use)[1]))
        
    for(hmo in names(hm_orders_to_use)){
      

        cat("#### ", paste0(hmo," {.tabset}"),"\n")
        cat("\n")
      # 
      #      cat("##### ", paste0("untrimmed"),"\n")
      # cat("\n")
      # 
      
      
      ### arrange rows by order to use
      curr_ann_row1 <- dplyr::arrange(annotation_row,  annotation_row[,hm_orders_to_use[[hmo]]]) 
      
    
      
      tmp2 <-  tmp[rownames(curr_ann_row1),intersect(rownames(annotation_col),colnames(tmp))] %>%
        ### remove columns with all NA
        dplyr::select(where(function(.) any(!is.na(.))))
      
      ### set the gaps for the rows from first element of orders, except if assay
      if(hm_orders_to_use[[hmo]][1] != "Assay"){
        gaps_df <- data.frame(table(curr_ann_row1[,hm_orders_to_use[[hmo]][1]])) %>%
          dplyr::mutate(cumsum = cumsum(Freq))
        if(length(unique(gaps_df$Freq)) ==1){
          if(unique(gaps_df$Freq) == 1){
            gaps_row_to_use <- 0
          }else{
            gaps_row_to_use <- unique(gaps_df$cumsum)
          }
          
        }else{
          gaps_row_to_use <- unique(gaps_df$cumsum)
        }
        
      }else{
        #gaps_df <- data.frame("cumsum"=c(0))
        gaps_row_to_use <- 0
      }
      
      #### gaps for column timepoints
      
      gaps_dfcol <- data.frame(table(annotation_col[colnames(tmp2),"Group_f"])) %>%
        dplyr::arrange(Var1) %>%
        dplyr::mutate(cumsum = cumsum(Freq))     
      
      
      
      
      ### remove variable so not plotted
      curr_ann_row <- curr_ann_row1%>%
        dplyr::select(-variable)
      



      
      
      annotation_col_for_plot <- annotation_col %>% 
        tibble::rownames_to_column() %>% 
        dplyr::filter(rowname %in% colnames(tmp2)) %>% 
        tibble::column_to_rownames() %>%
        dplyr::arrange("Group_f")
      
     ###### do you want to see the heatmap?
      
      if(plot_hm == "yes"){
      
      ############### don't plot this heatmap if amount since not helpful
        if(ds_to_use != "amount"){
      hc <-
        pheatmap::pheatmap(tmp2[,rownames(annotation_col_for_plot)],
                 scale=row_scale_set, 
                 na_col = "grey70",
                 cluster_cols=FALSE, 
                 cluster_rows = FALSE,
                 fontsize_row = row_font_size,
                 fontsize_col = 7,
                 gaps_col =unique(gaps_dfcol$cumsum),
                 gaps_row = gaps_row_to_use,
                 col=mycols, breaks=breakscale, 
                 clustering_method="ward.D2",
                 annotation_col = annotation_col_for_plot, 
                 annotation_row = curr_ann_row,
                 annotation_colors = annotation_colors)
      
      hc
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
        }else{
          ################ for amount make heatmap with deltas
          
          
                
  curr_phen_long1 <- phen_long_for_delta %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

  
vars_to_use1 <- unique(curr_phen_long1$variable)

  
  full_df1 <- curr_phen_long1 %>%
    ## trim to only Animal, var and delta
    dplyr::select(c(AnimalID,group_var_to_use,variable,delta)) %>%
    ## make wider
    tidyr::pivot_wider(names_from = "variable", values_from = "delta") %>%
    ## remove columns with only NA
    dplyr::select(where(function(x) !all(is.na(x)))) %>%
    

    dplyr::rename(Group_to_use=!!rlang::sym(group_var_to_use)) %>%

    ### factorize group
    dplyr::mutate(Group_f = factor(Group_to_use, intersect(
      names(group_colors_to_use),
      unique(data.frame(curr_phen_long1)[,group_var_to_use])))) %>%
    dplyr::select(-Group_to_use)
  
  
  annotation_col1 <- annotation_col %>%
    dplyr::filter(Group_f == data.frame(unique(curr_phen_long1[group_var_to_use]))[1,])
  
  
       tmp1 <- full_df1 %>%
        ### make rownames
        dplyr::mutate(animal_group = paste0(AnimalID,"__",as.character(Group_f))) %>%
      tibble::column_to_rownames("animal_group") %>%
        ##remove Group_f
        dplyr::select(-Group_f,-AnimalID) %>% 
        ## transpose to get columns = animals, rows = variables
        t() %>%
        data.frame() 
      
      tmp2_1 <-  tmp1[rownames(curr_ann_row1),intersect(rownames(annotation_col1),colnames(tmp1))] %>%
        ### remove columns with all NA
        dplyr::select(where(function(.) any(!is.na(.))))
      
      ### set the gaps for the rows from first element of orders, except if assay
      if(hm_orders_to_use[[hmo]][1] != "Assay"){
        gaps_df <- data.frame(table(curr_ann_row1[,hm_orders_to_use[[hmo]][1]])) %>%
          dplyr::mutate(cumsum = cumsum(Freq))
        if(length(unique(gaps_df$Freq)) ==1){
          if(unique(gaps_df$Freq) == 1){
            gaps_row_to_use <- 0
          }else{
            gaps_row_to_use <- unique(gaps_df$cumsum)
          }
          
        }else{
          gaps_row_to_use <- unique(gaps_df$cumsum)
        }
        
      }else{

        gaps_row_to_use <- 0
      }
      
      #### gaps for column timepoints
      
      gaps_dfcol1 <- data.frame(table(annotation_col1[colnames(tmp2_1),"Group_f"])) %>%
        dplyr::arrange(Var1) %>%
        dplyr::mutate(cumsum = cumsum(Freq))     
      
      
      
      
      ### remove variable so not plotted
      curr_ann_row_1 <- curr_ann_row1 %>%
        dplyr::select(-variable)
      



      
      
      annotation_col_for_plot1 <- annotation_col1 %>% 
        tibble::rownames_to_column() %>% 
        dplyr::filter(rowname %in% colnames(tmp2_1)) %>% 
        tibble::column_to_rownames() %>%
        dplyr::arrange("Group_f")
      
      
      hc1 <-
        pheatmap::pheatmap(tmp2_1[,rownames(annotation_col_for_plot1)],
                 scale="none", 
                 na_col = "grey70",
                 cluster_cols=FALSE, 
                 cluster_rows = FALSE,
                 fontsize_row = row_font_size,
                 fontsize_col = 7,
                 gaps_col =unique(gaps_dfcol1$cumsum),
                 gaps_row = gaps_row_to_use,
                 col=mycols, breaks=breakscale, 
                 clustering_method="ward.D2",
                 annotation_col = annotation_col_for_plot1, 
                 annotation_row = curr_ann_row_1,
                 annotation_colors = annotation_colors)
      
      hc1
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
      
      
      
      
      plotted_data_df <- rbind(plotted_data_df, tmp2_1[,rownames(annotation_col_for_plot1)] %>%
                                 ### add vs names(group_colors_to_use)[1]) to columnames
                                 `colnames<-`(paste0(colnames(.),paste0("vs",names(group_colors_to_use)[1]))) %>%
                                 tibble::rownames_to_column()

                                 )
      
      
      ####### and if cat_to_use_name == "AllVar" , also break out by category
      
      if(cat_to_use_name == "AllVar"){
        
        for(curr_cat_set in c("NK","DC","Monocyte")){
          
                           cat("##### ", paste0(curr_cat_set),"\n")
      cat("\n")
      
      
      curr_vars <-unique(phen_long_to_use$variable)[which(phen_long_to_use$category == curr_cat_set)]
      
      
       curr_ann_row1_trim <- curr_ann_row1 %>%
         dplyr::filter(variable %in% intersect(curr_ann_row1$variable,curr_vars)) 
       
       
         ### set the gaps for the rows from first element of orders, except if assay
      if(hm_orders_to_use[[hmo]][1] != "Assay"){
        gaps_dftrim <- data.frame(table(curr_ann_row1_trim[,hm_orders_to_use[[hmo]][1]])) %>%
          dplyr::mutate(cumsum = cumsum(Freq))
        if(length(unique(gaps_dftrim$Freq)) ==1){
          if(unique(gaps_dftrim$Freq) == 1){
            gaps_row_to_use_trim <- 0
          }else{
            gaps_row_to_use_trim <- unique(gaps_dftrim$cumsum)
          }
          
        }else{
          gaps_row_to_use_trim <- unique(gaps_dftrim$cumsum)
        }
        
      }else{

        gaps_row_to_use_trim <- 0
      }
      
      
     ### remove variable so not plotted
      curr_ann_row_1_trim <- curr_ann_row1_trim %>%
        dplyr::select(-variable)
      

      
       hc1 <-
        pheatmap::pheatmap(tmp2_1[rownames(curr_ann_row_1_trim),rownames(annotation_col_for_plot1)],
                 scale="none", 
                 na_col = "grey70",
                 cluster_cols=FALSE, 
                 cluster_rows = FALSE,
                 fontsize_row = row_font_size,
                 fontsize_col = 7,
                 gaps_col =unique(gaps_dfcol1$cumsum),
                 gaps_row = gaps_row_to_use_trim,
                 col=mycols, breaks=breakscale, 
                 clustering_method="ward.D2",
                 annotation_col = annotation_col_for_plot1, 
                 annotation_row = curr_ann_row_1_trim,
                 annotation_colors = annotation_colors)
      
      hc1
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      


          
        } ### end loop
        

        
        cat("\n")
      cat(' \n \n') ### this is the key!
      

  
        
        
        
        
        
      } ### end if all var

          
          
         cat("\n")
      cat(' \n \n') ### this is the key!
      

  
     
          
        }
      
          cat("\n")
      cat(' \n \n') ### this is the key!
      

      } ### end if want to see plot
      
          cat("\n")
      cat(' \n \n') ### this is the key!
      

  

      
    }
    
    cat("\n")
    cat(' \n \n') ### this is the key!
    
     list_of_plotted[[paste0(names_for_gtc)]] <- plotted_data_df
      
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
 
    
```

```{r scatterplots___NK_STIM, ref.label=c('scatterplots'), results='asis', eval=FALSE, fig.height=20, fig.width=7}
```





```{r setCat__Fig2F_DC}

cat_to_use <- "DC"
cat_to_use_name <- "Fig2F_DC"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes"

```

```{r setGroups___DC_amount, ref.label=c('setGroups___amount')}
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___DC_STIM, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=8}
```

```{r scatterplots___DC_STIM, ref.label=c('scatterplots'), results='asis', eval=FALSE, fig.height=16, fig.width=7}
```





```{r setCat__Fig2H_Mono}

cat_to_use <- "Monocyte"
cat_to_use_name <- "Fig2H_Mono"


### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
row_font_size <- 14
plot_hm <- "yes"

```

```{r setGroups___Monocytes_amount, ref.label=c('setGroups___amount')}
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___Mono_STIM, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=8}
```

```{r scatterplots___Mono_STIM, ref.label=c('scatterplots'), results='asis', eval=FALSE, fig.height=20, fig.width=7}
```

## Figure 3 {.tabset}

```{r setCat__Fig3A_NK}

cat_to_use <- "NK"
cat_to_use_name <- "Fig3A_NK"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes" ### plot the heatmap comparing the deltas?


```


```{r setGroups___amount_metA}

group_var_to_use <- "treatment"
group_colors_to_use <- c("GP120"="grey30",
                         "GP120_MetA"="deepskyblue"
                         )
names_for_gtc <- "MetA"

groups_to_compare <- data.frame("group1"=c("GP120"),                            
                                  "group2"=c("GP120_MetA")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)


#### for heatmap only
phen_long_for_delta <- phen_long

```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___metA___Fig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___metA___Fig3, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```




```{r setGroups___amount_SAMT}

group_var_to_use <- "treatment"
group_colors_to_use <- c("GP120"="grey30",
                         "GP120_samt247"="darkgreen"
                         )
names_for_gtc <- "SAMT"

groups_to_compare <- data.frame("group1"=c("GP120"),                            
                                  "group2"=c("GP120_samt247")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)


#### for heatmap only
phen_long_for_delta <- phen_long

```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___SAMT___Fig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___SAMT___Fig3, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```





```{r setGroups___amount_MetD}
group_var_to_use <- "treatment"
group_colors_to_use <- c("GP120"="grey30",
                         "GP120_MetD"="orange"
                         )
names_for_gtc <- "MetD"

groups_to_compare <- data.frame("group1"=c("GP120"),                            
                                  "group2"=c("GP120_MetD")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)


#### for heatmap only
phen_long_for_delta <- phen_long

```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___MetD___Fig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___MetD___Fig3, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```









```{r setPhen}

phen_long_to_use <- phen_long

### set dataset
ds_to_use <- c("delta") #c("amount","FC","delta")
row_scale_set <- "none"

    
           ### for delta
        mycols <- colorRampPalette(c("blue4","skyblue","white","lightpink","red4"))(1000)
        range <- 50
        breakscale <- c(-1*range,
                        seq(-.9*range,-.005*range, length.out=length(mycols)/4),
                        seq(-.0049*range,-.0001*range, length.out=length(mycols)/4-1),
                        0,
                        seq(.0001*range,.0049*range, length.out=length(mycols)/4-1),
                        seq(0.005*range,.9*range, length.out=length(mycols)/4),
                        1*range)
```



```{r setCat__Fig3BtoK_NK}

cat_to_use <- "NK"
cat_to_use_name <- "Fig3BtoK_NK"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "no" ### plot the heatmap comparing the deltas?


```





```{r setGroups_ASD}
group_var_to_use <- "treatment"
group_colors_to_use <- c(
                         "GP120_MetA"="deepskyblue",
                         "GP120_samt247"="darkgreen",
                        # "GP120_MetB"="black",
                        # "GP120_MetC"="blue",
                         "GP120_MetD"="orange"#,
                        # "GP120_NewDerivative"="purple3"
                         )
names_for_gtc <- "ASD"

groups_to_compare <- data.frame("group1"=c("GP120_samt247","GP120_samt247",
                                           "GP120_MetD"),                            
                                  "group2"=c("GP120_MetA","GP120_MetD",
                                            "GP120_MetA")) %>%
  dplyr::mutate(name = str_remove_all(str_remove_all(paste0(group2,"vs",group1),"GP120_"),"ivative"))


### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)
linewidth_to_use <- 0.5 ### non zero = yes add lines

```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups `r paste0(names_for_gtc)` 

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across 3 pairwise group contrasts*

comparing deltas = GP120_XXX - GP120

```{r diffBetweenGroups___ASD___Fig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=18}
```



```{r setVarsScatter_Fig3BCDEFGHIJK_NK}
cat_to_use_name <- "Fig3BCDEFGHIJK"

### sets order
vars_to_use <- c("Pct_NKp44_IL17","Pct_NK_CD73","Pct_NKp44_CD73","Pct_NK_CD107","Pct_NKp44",
                 "Pct_NK_TNFa","Pct_DN_IFNg","Pct_NK_IFNg","Pct_DN","Pct_NK"
)


  
  annotation_row <- annotation_row[vars_to_use,]
  
  
###### after seeing kmeans suggestions, give it better ones
  manual_ylim_ranges <- list(c(-2,2.0), c(-5, 5), c(-9,6), c(-25, 30))

```

```{r scatterplots___ASD___Fig3, ref.label=c('scatterplots'), results='asis', fig.height=16, fig.width=8}
```



## Supp Figure 1 {.tabset}

```{r setPhen_amount___SuppFig1, ref.label=c('setPhen_amount')}
```

```{r setCat__SuppFig1A_NK}

cat_to_use <- "NK"
cat_to_use_name <- "SuppFig1A_NK"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes" ### plot the heatmap comparing the deltas?


```


```{r setGroups___amount_metB}

group_var_to_use <- "treatment"
group_colors_to_use <- c("GP120"="grey30",
                         "GP120_MetB"="black"
                         )
names_for_gtc <- "MetB"

groups_to_compare <- data.frame("group1"=c("GP120"),                            
                                  "group2"=c("GP120_MetB")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)


#### for heatmap only
phen_long_for_delta <- phen_long
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___metB___SuppFig1, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___metB___SuppFig1, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```




```{r setGroups___amount_SAMT___SuppFig1, ref.label=c('setGroups___amount_SAMT')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___SAMT___SuppFig1, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___SAMT___SuppFig1, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```





```{r setGroups___amount_MetC}


group_var_to_use <- "treatment"
group_colors_to_use <- c("GP120"="grey30",
                         "GP120_MetC"="blue"
                         )
names_for_gtc <- "MetC"

groups_to_compare <- data.frame("group1"=c("GP120"),                            
                                  "group2"=c("GP120_MetC")) %>%
  dplyr::mutate(name = paste0(group2,"vs",group1))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)


#### for heatmap only
phen_long_for_delta <- phen_long
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___MetC___SuppFig1, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___MetC___SuppFig1, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```








```{r setPhen___SuppFig1, ref.label=c('setPhen')}
```



```{r setCat__SuppFig1BtoG_NK}

cat_to_use <- "NK"
cat_to_use_name <- "SuppFig1BtoG_NK"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "no" ### plot the heatmap comparing the deltas?


```





```{r setGroups_BSC}
group_var_to_use <- "treatment"
group_colors_to_use <- c("GP120_MetB"="black",
                         "GP120_samt247"="darkgreen",
                        # "GP120_MetA"="deepskyblue",
                         
                         "GP120_MetC"="blue"#,
                        # "GP120_MetD"="orange",
                        # "GP120_NewDerivative"="purple3"
                         )
names_for_gtc <- "BSC"

groups_to_compare <- data.frame("group1"=c("GP120_samt247","GP120_samt247",
                                           "GP120_MetC"),                            
                                  "group2"=c("GP120_MetB","GP120_MetC",
                                            "GP120_MetB")) %>%
  dplyr::mutate(name = str_remove_all(str_remove_all(paste0(group2,"vs",group1),"GP120_"),"ivative"))

### trim to only the variables for this fig
  curr_phen_long <- phen_long_to_use %>%
    dplyr::filter(category %in% cat_to_use) %>%
    dplyr::filter(treatment %in% names(group_colors_to_use))

vars_to_use <- unique(curr_phen_long$variable)
linewidth_to_use <- 0.5 ### non zero = yes add lines
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups `r paste0(names_for_gtc)` 

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across 3 pairwise group contrasts*

comparing deltas = GP120_XXX - GP120

```{r diffBetweenGroups___BSC___SuppFig1, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=18}
```



```{r setVarsScatter_SuppFig1BCDEFG_NK}
cat_to_use_name <- "SuppFig1BCDEFG"

### sets order
vars_to_use <- c("Pct_NKp44_CD73","Pct_NK_CD73","Pct_DN_CD73","Pct_NK_CD107",
                 "Pct_NKp44","Pct_DN"
)

### trim to only the variables for this fig
  annotation_row <- annotation_row[vars_to_use,]

```

```{r scatterplots___BSC___SuppFig1, ref.label=c('scatterplots'), results='asis', fig.height=12, fig.width=8}
```







## Figure 4 {.tabset}

```{r setPhen_amount___Fig4, ref.label=c('setPhen_amount')}
```

```{r setCat__Fig4A_DC}

cat_to_use <- "DC"
cat_to_use_name <- "Fig4A_DC"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes" ### plot the heatmap comparing the deltas?


```


```{r setGroups___amount_metA___Fig4, ref.label=c('setGroups___amount_metA')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___metA___Fig4, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___metA___Fig4, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```




```{r setGroups___amount_SAMT___Fig4, ref.label=c('setGroups___amount_SAMT')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___SAMT___Fig4, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___SAMT___Fig4, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```





```{r setGroups___amount_MetD___Fig4, ref.label=c('setGroups___amount_MetD')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___MetD___Fig4, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___MetD___Fig4, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```








```{r setPhen___Fig4, ref.label=c('setPhen')}
```



```{r setCat__Fig4BtoI_DC}

cat_to_use <- "DC"
cat_to_use_name <- "Fig4BtoI_DC"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "no" ### plot the heatmap comparing the deltas?


```





```{r setGroups_ASD___Fig4, ref.label=c('setGroups_ASD')}
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups `r paste0(names_for_gtc)` 

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across 3 pairwise group contrasts*

comparing deltas = GP120_XXX - GP120

```{r diffBetweenGroups___ASD___Fig4, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=18}
```



```{r setVarsScatter_Fig4BCDEFGHI_DC}
cat_to_use_name <- "Fig4BCDEFGHI"

### sets order
vars_to_use <- c("Pct_DC10_IL10","Pct_mDC_IL10","Pct_mDC_IFNg","Pct_DC10_CD73","Pct_mDC_CD73",
                 "Pct_DC10_TNFa","Pct_mDC_TNFa","Pct_DC10"
)

### trim to only the variables for this fig
  annotation_row <- annotation_row[vars_to_use,]
  
  
  ###### after seeing kmeans suggestions, give it better ones
  manual_ylim_ranges <- list(c(-1.0,0.6), c(-5, 4), c(-20,20), c(-25, 40))

```

```{r scatterplots___ASD___Fig4, ref.label=c('scatterplots'), results='asis', fig.height=12, fig.width=8}
```






## Supp Figure 2 {.tabset}

```{r setPhen_amount___SuppFig2, ref.label=c('setPhen_amount')}
```

```{r setCat__SuppFig2A_DC}

cat_to_use <- "DC"
cat_to_use_name <- "SuppFig2A_DC"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes" ### plot the heatmap comparing the deltas?


```


```{r setGroups___amount_metB___SuppFig2, ref.label=c('setGroups___amount_metB')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___metB___SuppFig2, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___metB___SuppFig2, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```




```{r setGroups___amount_SAMT___SuppFig2, ref.label=c('setGroups___amount_SAMT')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___SAMT___SuppFig2, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___SAMT___SuppFig2, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```





```{r setGroups___amount_MetC___SuppFig2, ref.label=c('setGroups___amount_MetC')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___MetC___SuppFig2, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___MetC___SuppFig2, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```








```{r setPhen___SuppFig2, ref.label=c('setPhen')}
```



```{r setCat__SuppFig2BtoF_DC}

cat_to_use <- "DC"
cat_to_use_name <- "SuppFig2BtoF_DC"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "no" ### plot the heatmap comparing the deltas?


```





```{r setGroups_BSC___SuppFig2, ref.label=c('setGroups_BSC')}
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups `r paste0(names_for_gtc)` 

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across 3 pairwise group contrasts*

comparing deltas = GP120_XXX - GP120

```{r diffBetweenGroups___BSC___SuppFig2, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=18}
```



```{r setVarsScatter_SuppFig2BCDEFG_DC}
cat_to_use_name <- "SuppFig2BCDEFG"

### sets order
vars_to_use <- c("Pct_DC10_IL10","Pct_DC10_CD73",#"Pct_mDC_CD73",
                 "Pct_mDC_IL10","Pct_DC10","Pct_mDC_TNFa")

### trim to only the variables for this fig
  annotation_row <- annotation_row[vars_to_use,]

```

```{r scatterplots___BSC___SuppFig2, ref.label=c('scatterplots'), results='asis', fig.height=8, fig.width=8}
```










## Figure 5 {.tabset}

```{r setPhen_amount___Fig5, ref.label=c('setPhen_amount')}
```

```{r setCat__Fig5A_Mono}

cat_to_use <- "Monocyte"
cat_to_use_name <- "Fig5A_Mono"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes" ### plot the heatmap comparing the deltas?


```


```{r setGroups___amount_metA___Fig5, ref.label=c('setGroups___amount_metA')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___metA___Fig5, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___metA___Fig5, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```




```{r setGroups___amount_SAMT___Fig5, ref.label=c('setGroups___amount_SAMT')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___SAMT___Fig5, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___SAMT___Fig5, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```





```{r setGroups___amount_MetD___Fig5, ref.label=c('setGroups___amount_MetD')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___MetD___Fig5, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___MetD___Fig5, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```








```{r setPhen___Fig5, ref.label=c('setPhen')}
```



```{r setCat__Fig5BtoI_Mono}

cat_to_use <- "Monocyte"
cat_to_use_name <- "Fig5BtoI_Mono"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "no" ### plot the heatmap comparing the deltas?


```





```{r setGroups_ASD___Fig5, ref.label=c('setGroups_ASD')}
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups `r paste0(names_for_gtc)` 

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across 3 pairwise group contrasts*

comparing deltas = GP120_XXX - GP120

```{r diffBetweenGroups___ASD___Fig5, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=18}
```



```{r setVarsScatter_Fig5BCDEFGH_Mono}
cat_to_use_name <- "Fig5BCDEFGH"

### sets order
vars_to_use <- c("Pct_Classical_CD73","Pct_Classical_IL10","Pct_Nonclassical_CD73","Pct_Nonclassical_IFNg",
                 "Pct_Classical","Pct_Intermediate","Pct_Nonclassical"
)

### trim to only the variables for this fig
  annotation_row <- annotation_row[vars_to_use,]

```

```{r scatterplots___ASD___Fig5, ref.label=c('scatterplots'), results='asis', fig.height=12, fig.width=8}
```




## Supp Figure 3 {.tabset}

```{r setPhen_amount___SuppFig3, ref.label=c('setPhen_amount')}
```

```{r setCat__SuppFig3A_Mono}

cat_to_use <- "Monocyte"
cat_to_use_name <- "SuppFig3A_Mono"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "yes" ### plot the heatmap comparing the deltas?


```


```{r setGroups___amount_metB___SuppFig3, ref.label=c('setGroups___amount_metB')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___metB___SuppFig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___metB___SuppFig3, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```




```{r setGroups___amount_SAMT___SuppFig3, ref.label=c('setGroups___amount_SAMT')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___SAMT___SuppFig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___SAMT___SuppFig3, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```





```{r setGroups___amount_MetC___SuppFig3, ref.label=c('setGroups___amount_MetC')}
```




### `r paste0(cat_to_use_name, " " ,names_for_gtc)` Differences vs GP120  {.tabset}

*paired MannWhitney/Wilcoxon, raw p values used since pair of groups considered*

comparing raw values, heatmaps are deltas


```{r diffBetweenGroups___MetC___SuppFig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=7}
```

```{r scatterplots___MetC___SuppFig3, ref.label=c('scatterplots'), eval=FALSE, results='asis', fig.height=25, fig.width=7}
```








```{r setPhen___SuppFig3, ref.label=c('setPhen')}
```



```{r setCat__SuppFig3BCD_Mono}

cat_to_use <- "Monocyte"
cat_to_use_name <- "SuppFig3BCD_Mono"

### scatterplot for non-sig variables too?
plot_not_sig <- "yes"
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
plot_hm <- "no" ### plot the heatmap comparing the deltas?


```





```{r setGroups_BSC___SuppFig3, ref.label=c('setGroups_BSC')}
```

### `r paste0(cat_to_use_name)` {.tabset}

Differences between groups `r paste0(names_for_gtc)` 

*paired MannWhitney/Wilcoxon, Benjamini-Hochberg adjusted for multiple comparisons across 3 pairwise group contrasts*

comparing deltas = GP120_XXX - GP120

```{r diffBetweenGroups___BSC___SuppFig3, ref.label=c('diffBetweenGroups'), results='asis', fig.height=8, fig.width=18}
```



```{r setVarsScatter_SuppFig3BCD_Mono}
cat_to_use_name <- "SuppFig3BCD"

### sets order
vars_to_use <- c("Pct_Classical_CD73","Pct_Classical_IL10","Pct_Nonclassical_IFNg")

### trim to only the variables for this fig
  annotation_row <- annotation_row[vars_to_use,]
  
  n_ylim_buckets <- 2   ##only 3 vars in this figure, use 2 ranges not 3


```

```{r scatterplots___BSC___SuppFig3, ref.label=c('scatterplots'), results='asis', fig.height=8, fig.width=8}
```








## Figure 6 / Supp Figure 4 {.tabset}

```{r setVsVar__SuppFig4_Fig6}
  vsVar_to_use <- "MetA_Week_15_minusAverage"

```

### `r paste0(vsVar_to_use)` {.tabset}

#### Phenotype {.tabset}

```{r Fig6phen_prep}
group_var_to_use <- "Derivative"
group_colors_to_use <- c("MetA"="deepskyblue"
                         )
names_for_gtc <- "Fig6_phen"
cat_to_use_name <- names_for_gtc
curr_ds <- "amount"

fig6_manual_ranges <- list(c(0.0, 0.5), c(0,60))   ### tune these to your phenotype data


### trim to only the variables for this fig
  curr_phen_long <- rawdatasets_list[["Study-2"]] %>% 
    ### make long
    tidyr::pivot_longer(!c(AnimalID,TOA,MetA_in_plasma__Baseline,MetA_in_plasma__Week15,MetA_Week_15_minusAverage), names_to = "variable", values_to = "amount") %>% 
    dplyr::filter(grepl("Phenotype",variable))
  
  
  
  curr_df_wide <- curr_phen_long %>%
dplyr::select(AnimalID,variable,amount,!!rlang::sym(vsVar_to_use)) %>%
    tidyr::pivot_wider(names_from = "variable",values_from = "amount") %>%
    column_to_rownames("AnimalID")
  

vars_to_use_prep <- unique(curr_phen_long$variable)
### set order
vars_to_use <- c(
  vars_to_use_prep[grepl(paste(c("NK","DN"), collapse = '|'),vars_to_use_prep)],
  vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)][!grepl("_",vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)])],
  vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)][grepl("_",vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)])],
  
  vars_to_use_prep[grepl(paste(c("Macro","DC"), collapse = '|'),vars_to_use_prep)][!grepl("_",vars_to_use_prep[grepl(paste(c("Macro","DC"), collapse = '|'),vars_to_use_prep)])],
  vars_to_use_prep[grepl(paste(c("Macro","DC"), collapse = '|'),vars_to_use_prep)][grepl("_",vars_to_use_prep[grepl(paste(c("Macro","DC"), collapse = '|'),vars_to_use_prep)])]
  
)




linewidth_to_use <- 0 ### zero = no lines, non zero = yes add lines


### scatterplot for non-sig variables too?
plot_not_sig <- "no"
p_cut_for_scatterplots <- 0.1
num_columns_for_plot <- 3
cex_for_beeswarm <- 2
y_axis_size <- 10 
row_font_size <- 14




shape_values_to_use <-c(0,1,2,5,6,8,12,21,22,23,24,25)

```


```{r AssocWithMetAPlasma, results='asis', fig.height=4, fig.width=13}

  
  cor_method <- "spearman"
  
  ## full_df only has the variables in columns and the factorized group (plus any other covariates if they are in the model)
  

  
   curr_df1 <- curr_df_wide %>%
     dplyr::select(!!!rlang::syms(vars_to_use))
   
   
      curr_df2 <- curr_df_wide %>%
        ## trim to only the variable columns
        dplyr::select(!!rlang::sym(vsVar_to_use)) 
      
      ########## do correlation with psych package since much faster
      sp <- psych::corr.test(as.matrix(signif(curr_df1,7)), as.matrix(signif(curr_df2,7)), method = cor_method,
                             use="pairwise",
                             adjust="none",ci=FALSE)
      
      cor_mat <- sp$r
      cor_test_mat <- data.frame(sp$p)
      
      ## adj p for multiple comparisons by BH over assays
      cor_test_adj_mat <- cor_test_mat #duplicate dataframe to replace values
      for(curr_assay in colnames(curr_df1)) {
        cor_test_adj_mat[curr_assay,] <- p.adjust(cor_test_mat[curr_assay,], method = "BH")
      }
      
      
      #### also combine to dataframe since only one columns
      corr_df <- dplyr::left_join(data.frame(cor_mat) %>%
                                    rownames_to_column() %>%
                                    dplyr::rename(R_value=2),
                                  data.frame(cor_test_mat) %>%
                                    rownames_to_column() %>%
                                    dplyr::rename(raw_p=2),
                                  by = "rowname") %>%
        dplyr::left_join(data.frame(cor_test_adj_mat) %>%
                           rownames_to_column() %>%
                           dplyr::rename(adj_p=2),
                         by = "rowname") %>%
        ### add columns
        dplyr::mutate(
          ### set R_value to NA if p is NA i.e there were only 2 datapoints, some how the R is calculated but the p is not
          R_value = ifelse(is.na(raw_p),NA,R_value),
          dir = ifelse(R_value > 0 & raw_p < 0.05,
                       "up", 
                       ifelse(R_value < 0 & raw_p < 0.05,
                              "down","not_sig"))) 
      
      
      ### save stats
    list_of_stats[[paste0(names_for_gtc,"_",cor_method)]] <- corr_df
    
    # saveRDS(corr_df_combo, file = file.path(outputRData_dir, 
    #                                                               paste0(Corr_name,"_", cor_method,"_CoefPvals_untrimmed.rds")))
    
    
    cat("\n")
    cat(' \n \n') ### this is the key!
    
    
    ################## now plot sig correlations for all groups
    
    
    
    
    if(grepl("phen",names_for_gtc)){

      #### set low/high range
      low_range <- data.frame(rowname="range_set_low",
                              bulk=as.numeric(1),
                              CD73pos=as.numeric(1))
      
      
      high_range <- data.frame(rowname="range_set_high",
                               bulk=as.numeric(1),
                               CD73pos=as.numeric(1))
      
      
      
      
      low_Rrange <- data.frame(rowname="range_set_low",
                               bulk=as.numeric(-1),
                               CD73pos=as.numeric(-1))
      
      
      high_Rrange <- data.frame(rowname="range_set_high",
                                bulk=as.numeric(1),
                                CD73pos=as.numeric(1))
   
    
    
    
    cor_test_mat_for_plot <- corr_df %>%
      ### trim to only p values
      dplyr::select(rowname,raw_p) %>%
      ## separate CD73 & bulk
      dplyr::mutate(marker = ifelse(grepl("CD73pos",rowname),"CD73pos","bulk"),
                    rowname = str_remove_all(rowname,"CD73pos_"),
                    rowname = str_remove_all(rowname,"_"),
                    raw_p = round(raw_p, 3)) %>%
      ## make wider
      tidyr::pivot_wider(names_from = "marker",values_from = "raw_p") %>%
      
      column_to_rownames() %>%
      ## filter to only vars that are sig in at least one group
   #   dplyr::filter_at(vars(everything(.)), any_vars(. < 0.1)) %>%
      rownames_to_column() %>%
      #### add dummy row to set square size range
      add_row(low_range) %>%
      add_row(high_range) %>%
      column_to_rownames() %>%
      as.matrix() 
    
    cor_mat_for_plot <- corr_df %>%
      ### trim to only R values
      dplyr::select(rowname,R_value) %>%
            ## separate CD73 & bulk
      dplyr::mutate(marker = ifelse(grepl("CD73pos",rowname),"CD73pos","bulk"),
                    rowname = str_remove_all(rowname,"CD73pos_"),
                    rowname = str_remove_all(rowname,"_")) %>%
      ## make wider
      tidyr::pivot_wider(names_from = "marker",values_from = "R_value") %>%
      #### add dummy row to set square size range
      add_row(low_Rrange) %>%
      add_row(high_Rrange) %>%
      column_to_rownames() %>%
      as.matrix()
    
     ann_pval_level <- 0.1
    tl_cex <- 1
    cl_ratio <- 1.2
    pch_cex <- 3
    
   
    
    }else{
      
      
      
      
        #### set low/high range
      low_range <- data.frame(rowname="range_set_low",
                              gp120=as.numeric(1),
                              PMA=as.numeric(1))
      
      
      high_range <- data.frame(rowname="range_set_high",
                               gp120=as.numeric(1),
                               PMA=as.numeric(1))
      
      
      
      
      low_Rrange <- data.frame(rowname="range_set_low",
                               gp120=as.numeric(-1),
                               PMA=as.numeric(-1))
      
      
      high_Rrange <- data.frame(rowname="range_set_high",
                                gp120=as.numeric(1),
                                PMA=as.numeric(1))
   
    
    
    
    cor_test_mat_for_plot <- corr_df %>%
      ### trim to only p values
      dplyr::select(rowname,raw_p) %>%
      ## separate CD73 & gp120
      dplyr::mutate(marker = ifelse(grepl("PMA",rowname),"PMA","gp120"),
                    rowname = str_after_first(rowname,"___"),
                    raw_p = round(raw_p, 3)) %>%
      ## make wider
      tidyr::pivot_wider(names_from = "marker",values_from = "raw_p") %>%
      
      column_to_rownames() %>%
      ## filter to only vars that are sig in at least one group
   #   dplyr::filter_at(vars(everything(.)), any_vars(. < 0.1)) %>%
      rownames_to_column() %>%
      #### add dummy row to set square size range
      add_row(low_range) %>%
      add_row(high_range) %>%
      column_to_rownames() %>%
      as.matrix() 
    
    cor_mat_for_plot <- corr_df %>%
      ### trim to only R values
      dplyr::select(rowname,R_value) %>%
            ## separate CD73 & gp120
      dplyr::mutate(marker = ifelse(grepl("PMA",rowname),"PMA","gp120"),
                    rowname = str_after_first(rowname,"___")) %>%
      ## make wider
      tidyr::pivot_wider(names_from = "marker",values_from = "R_value") %>%
      #### add dummy row to set square size range
      add_row(low_Rrange) %>%
      add_row(high_Rrange) %>%
      column_to_rownames() %>%
      as.matrix()
    
     ann_pval_level <- 0.05
    tl_cex <- 1
    cl_ratio <- 1.2
    pch_cex <- 3
    

      
      
    }
    
    
       
      cat("##### ",paste0(" corrplot"),"\n")
      cat('\n')
      
         cat(" ",paste0(" stars for p < ",ann_pval_level),"\n")
      cat('\n')
  
      
      #### otherwise can use the stars
      corrplot::corrplot(cor_mat_for_plot, method = 'square', 
                         is.corr = FALSE, #since not symmetric
                         p.mat = cor_test_mat_for_plot,
                         sig.level = as.numeric(ann_pval_level), 
                         insig='label_sig',  #set p < 0.05 to blank
                         pch.cex = pch_cex, pch.col = 'white',tl.col = 'black',tl.cex=tl_cex,
                         ##legend
                         cl.ratio = cl_ratio, #cl.cex = 2,
                         #addCoef.col ='grey80', 
                         #number.cex = 0.5, #add corr coeff to sig circles
                         #order = order, ## CANNOT ORDER ASSYMMETRIC
                         col = colorRampPalette(c("turquoise4","white","darkorange"))(200), ### change colors so don't match DNA
                         # col = (COL2('PuOr',1000)), ##make orange NEGATIVE
                         col.lim = c(-1,1) ##set limits to be the same across plots
                         # title = vaccine_groups[i], mar=c(0,0,1,0) # http://stackoverflow.com/a/14754408/54964
      )
      
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
      
      cat("##### ",paste0(" scatterplots "),"\n")
      cat('\n')
  
######## collect data to plot, make vsVar_to_use a separate column from variable
      df_to_plot <- curr_df_wide %>%
        rownames_to_column("AnimalID") %>%
        tidyr::pivot_longer(!c(AnimalID,!!rlang::sym(vsVar_to_use)),
                            names_to = "variable",
                            values_to = "amount")
      
      ### moved up from below - need vars_to_use1 before computing y-axis ranges
      if(plot_not_sig == "yes"){
        vars_to_use1 <- vars_to_use
      }else{
      vars_to_use1 <- corr_df %>%
        dplyr::filter(raw_p < p_cut_for_scatterplots) %>%   
        dplyr::pull(rowname) %>%
        intersect(vars_to_use)
      }
      
      ### -------- y-axis range control for Fig6 scatter/correlation plots --------
      ### ranges are set manually per prep chunk (fig6_manual_ranges), not clustered.
      ### each prep chunk (Fig6phen_prep / Fig6stim_prep) should define e.g.:
      ###   fig6_manual_ranges <- list(c(-1, 1), c(-10, 10))
      ### ranges should be listed smallest to largest.
      if (exists("fig6_manual_ranges")) {
        candidate_ranges_fig6 <- fig6_manual_ranges
      } else {
        ### fallback if a prep chunk forgets to set it - one wide range
        candidate_ranges_fig6 <- list(c(-10, 10))
      }
      
      print(candidate_ranges_fig6)
      ### -------- end y-axis range control --------
      
      list_of_plots <- list()
      plotted_data_df <- data.frame()
      
      
  
      
   for (i in vars_to_use1){
        #### trim to only this variable
        curr_df_to_plot <- df_to_plot %>%
          dplyr::filter(variable == i)
        
        ### pick smallest bucket containing this variable's y-data
        data_min_fig6 <- min(curr_df_to_plot$amount, na.rm = TRUE)
        data_max_fig6 <- max(curr_df_to_plot$amount, na.rm = TRUE)
        
        range_to_use_fig6 <- candidate_ranges_fig6[[length(candidate_ranges_fig6)]]
        for (r in candidate_ranges_fig6) {
          if (data_min_fig6 >= r[1] && data_max_fig6 <= r[2]) {
            range_to_use_fig6 <- r
            break
          }
        }
        
        breaks_fig6 <- get_zero_anchored_breaks(range_to_use_fig6[1], range_to_use_fig6[2])
        range_to_use_fig6 <- c(min(range_to_use_fig6[1], min(breaks_fig6)),
                                max(range_to_use_fig6[2], max(breaks_fig6)))
        

          p <-  ggplot(curr_df_to_plot, aes(y = amount, 
                                            x = !!rlang::sym(vsVar_to_use)))+
            
            coord_cartesian(ylim = range_to_use_fig6) +
            scale_y_continuous(breaks = breaks_fig6) +
            geom_smooth(se = FALSE, method = "lm", color = "black")+
            
            geom_point(aes(shape = AnimalID), 
                       color = unname(group_colors_to_use), fill = unname(group_colors_to_use),
                       size = 5, stroke = 1, alpha = 0.8, cex = 3)+
            scale_shape_manual(values = shape_values_to_use)+
            
            
            ylab(paste0(i))+
            labs(title = paste0("R = ",signif(as.double(corr_df[which(corr_df$rowname == i),"R_value"]),2)),
                 subtitle =paste0("p = ",signif(as.double(corr_df[which(corr_df$rowname == i),"raw_p"]),2)))+
            theme_bw()+
            #### change color in legend for fill to have shape that can fill
           # guides(fill = guide_legend(override.aes = list(shape = 21)))+
            
            theme(legend.position="right", 
                  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                  panel.border = element_blank(), 
                  axis.line = element_line(colour = "black",lineend = "square"),
                  axis.title = element_text(size = 15),
                  axis.text = element_text(size = 15),
                  plot.title =  element_text(size = 15, color = group_colors_to_use), plot.subtitle = element_text(size = 15, color = group_colors_to_use),
            )
          
       
        p1 <- p
        list_of_plots[[i]] <- p1 + theme(legend.position="none")
        
        
        
        cat("\n")
        cat(' \n \n') ### this is the key!
        
        plotted_data_df <- rbind(plotted_data_df, curr_df_to_plot)
        
      }
      
      ### get the legend
      #https://stackoverflow.com/questions/13649473/add-a-common-legend-for-combined-ggplots
      g_legend<-function(a.gplot){
        tmp <- ggplot_gtable(ggplot_build(a.gplot))
        leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
        legend <- tmp$grobs[[leg]]
        return(legend)}
      
      ### empty filler
      # list_of_plots[[nrow(list_of_subsets_for_plotting[[paste0(set)]])+1]]<-ggplot() + theme_void()
      # list_of_plots[[nrow(list_of_subsets_for_plotting[[paste0(set)]])+2]]<-g_legend(p1) ### grabs from the last plot
      list_of_plots[[length(list_of_plots)+1]]<-g_legend(p1) ### grabs from the last plot
       
      do.call("grid.arrange",c(list_of_plots, ncol=num_columns_for_plot))
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
      list_of_plotted[[paste0(names_for_gtc)]] <- plotted_data_df
      
      
      
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
      
      
 


```




#### gp120 & PMA stim {.tabset}

```{r Fig6stim_prep}
group_var_to_use <- "Derivative"
group_colors_to_use <- c("MetA"="deepskyblue"
                         )
names_for_gtc <- "Fig6_stim"
cat_to_use_name <- names_for_gtc
curr_ds <- "amount"

fig6_manual_ranges <- list(c(0, 6), c(0,20))   ### tune these to your phenotype data




### trim to only the variables for this fig
  curr_phen_long <- rawdatasets_list[["Study-2"]] %>% 
    ### make long
    tidyr::pivot_longer(!c(AnimalID,TOA,MetA_in_plasma__Baseline,MetA_in_plasma__Week15,MetA_Week_15_minusAverage), names_to = "variable", values_to = "amount") %>% 
    dplyr::filter(!grepl("Phenotype",variable))
  
  
  
  curr_df_wide <- curr_phen_long %>%
dplyr::select(AnimalID,variable,amount,!!rlang::sym(vsVar_to_use)) %>%
    tidyr::pivot_wider(names_from = "variable",values_from = "amount") %>%
    column_to_rownames("AnimalID")
  
  

vars_to_use_prep <- unique(curr_phen_long$variable)
### set order
vars_to_use <- c(

  
  vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)][grepl("gp120",vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)])],
  vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)][grepl("PMA",vars_to_use_prep[grepl(paste(c("lassical","Inter"), collapse = '|'),vars_to_use_prep)])],
  
  vars_to_use_prep[grepl(paste(c("MQ","DC"), collapse = '|'),vars_to_use_prep)][grepl("gp120",vars_to_use_prep[grepl(paste(c("MQ","DC"), collapse = '|'),vars_to_use_prep)])],
  vars_to_use_prep[grepl(paste(c("MQ","DC"), collapse = '|'),vars_to_use_prep)][grepl("PMA",vars_to_use_prep[grepl(paste(c("MQ","DC"), collapse = '|'),vars_to_use_prep)])], 
  
  
    vars_to_use_prep[grepl(paste(c("NK","DN"), collapse = '|'),vars_to_use_prep)][grepl("gp120",vars_to_use_prep[grepl(paste(c("NK","DN"), collapse = '|'),vars_to_use_prep)])],
  vars_to_use_prep[grepl(paste(c("NK","DN"), collapse = '|'),vars_to_use_prep)][grepl("PMA",vars_to_use_prep[grepl(paste(c("NK","DN"), collapse = '|'),vars_to_use_prep)])]
  
)




linewidth_to_use <- 0 ### zero = no lines, non zero = yes add lines


### scatterplot for non-sig variables too?
plot_not_sig <- "no"
p_cut_for_scatterplots <- 0.1
num_columns_for_plot <- 5
cex_for_beeswarm <- 2
y_axis_size <- 10 
row_font_size <- 14




shape_values_to_use <-c(0,1,2,5,6,8,12,21,22,23,24,25)

```


```{r AssocWithMetAPlasma___stim, ref.label=c('AssocWithMetAPlasma'), results='asis', fig.height=8, fig.width=21.5}
```


































## Figure 7 {.tabset}


```{r fig7summary, results='asis', eval=TRUE}
#list_of_stats <- readRDS(file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_stats.rds")))
fig_name <- "Fig7"

### collect vs SAMT response deltas
vsSAMTres_df <- data.frame()
for(curr_tab in c(names(list_of_stats)[grepl(paste0(c("Fig3Bto","Fig4Bto","Fig5Bto","SuppFig1Bto","SuppFig2Bto","SuppFig3BCD_Mono"), collapse = '|'),names(list_of_stats))])){
  
  
  curr_df <- list_of_stats[[paste0(curr_tab)]] %>%
    ## collect dir column
    dplyr::select(variable, ends_with("vssamt247__dir")) %>%
    ##make longer
    tidyr::pivot_longer(!variable, names_to = "Group", values_to = "dir") %>%
    dplyr::mutate(Group = str_remove_all(Group,"__dir"))
  vsSAMTres_df <- rbind(vsSAMTres_df, curr_df)
  
}

vsSAMTres_df_wide <- vsSAMTres_df %>%
  ### remove dups for samt in both main and supp figs
  unique() %>%
  tidyr::pivot_wider(names_from = "Group", values_from = "dir") %>%
  ## bring samt to left & sort by it
  dplyr::relocate(variable,sort(colnames(.)[grepl("Met",colnames(.))])) %>%
  dplyr::arrange(!!!rlang::syms(setdiff(colnames(.),"variable"))) 


### collect vs gp120
vsGP120_df <- data.frame()
for(curr_tab in c(names(list_of_stats)[grepl(paste0(c("Fig3A_","Fig4A_","Fig5A_","SuppFig1A_","SuppFig2A_"), collapse = '|'),names(list_of_stats))])){
  curr_df <- list_of_stats[[paste0(curr_tab)]] %>%
    ## collect dir column
    dplyr::select(rowname, ends_with("dir")) %>%
    ##make longer
    tidyr::pivot_longer(!rowname, names_to = "Group", values_to = "dir") %>%
    dplyr::mutate(Group = str_remove_all(Group,"__dir"))
  vsGP120_df <- rbind(vsGP120_df, curr_df)
}

vsGP120_df_wide <- vsGP120_df %>%
  ### remove dups for samt in both main and supp figs
  unique() %>%
  tidyr::pivot_wider(names_from = "Group", values_from = "dir") %>%
  ## bring samt to left & sort by it
  dplyr::relocate(rowname,GP120_samt247vsGP120,sort(colnames(.)[grepl("Met",colnames(.))])) %>%
  dplyr::arrange(!!!rlang::syms(setdiff(colnames(.),"rowname"))) %>%
  dplyr::rename(variable=rowname) %>%
  ### add the vsSAMT response dirs
  dplyr::left_join(vsSAMTres_df_wide, by = "variable") %>%
  ### convert p < 0.1 to not_sig
  # dplyr::mutate(across(
  #   -variable,
  #   ~ factor(
  #       case_when(
  #         as.character(.x) %in% c("up, p < 0.1", "down, p < 0.1") ~ "not_sig",
  #         TRUE ~ as.character(.x)
  #       ),
  #       levels = c("up, p < 0.05", "not_sig", "down, p < 0.05")
  #     )
  # )) %>%
  dplyr::arrange(!!!rlang::syms(setdiff(colnames(.),"variable"))) 

list_of_plotted[[fig_name]] <- vsGP120_df_wide
list_of_stats[[fig_name]] <- vsGP120_df_wide

  to_print <- vsGP120_df_wide %>%
    DT::datatable(extensions = 'Buttons',
                              options = list(dom = 'Blfrtip',
                                             buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                             iDisplayLength = 50,
                                             lengthMenu = list(c(10,25,50,-1),
                                                               c(10,25,50,"All"))))
              
              
              print(htmltools::tagList(to_print))


       cat("\n")
        cat(' \n \n') ### this is the key!
  


```







## Supp Figure 5 {.tabset}

BH adjusted (over 3 pairwise between group comparisons) p for differences between groups




```{r fig5summary, results='asis', eval=TRUE}
#list_of_stats <- readRDS(file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_stats.rds")))
fig_name <- "SuppFig5"



### collect gp120 vsunstim
vsUNSTIM_df <- data.frame()
for(curr_tab in c(names(list_of_stats)[grepl(paste0(c("Fig2C_","Fig2F_","Fig2H_"), collapse = '|'),names(list_of_stats))])){
  
  
  curr_df <- list_of_stats[[paste0(curr_tab)]] %>%
    ## collect dir column
    dplyr::select(variable, ends_with("__dir")) %>%
    ##make longer
    tidyr::pivot_longer(!variable, names_to = "Group", values_to = "dir") %>%
    dplyr::mutate(Group = str_remove_all(Group,"__dir"))
  vsUNSTIM_df <- rbind(vsUNSTIM_df, curr_df)
  
}

vsUNSTIM_df_wide <- vsUNSTIM_df %>%
  tidyr::pivot_wider(names_from = "Group", values_from = "dir") 


##### add to Fig7
sfig5_df <- list_of_stats[["Fig7"]] %>%
  dplyr::left_join(vsUNSTIM_df_wide, by = "variable") %>%
  dplyr::relocate(variable,GP120vsUNSTIM) %>%
  ## sort by variable
  dplyr::mutate(var_only = ifelse(str_count(variable,"_") == 1,
                                  variable,
                                  str_before_last(variable,"_")),
                var_only = factor(var_only, paste0("Pct_",c("DN","NKp44","NK","mDC","DC10","Classical","Intermediate","Nonclassical"))),
                marker = ifelse(str_count(variable,"_") == 1,
                                  paste0("total"),
                                  str_after_last(variable,"_")),
                marker = factor(marker, c("total","IL17","CD107","CD73","IL10","IFNg","TNFa"))) %>%
  dplyr::arrange(marker, var_only)



list_of_plotted[[fig_name]] <- sfig5_df
list_of_stats[[fig_name]] <- sfig5_df
```

```{r suppfig5generate, fig.height = 12, eval=TRUE}
df <- sfig5_df

## ---------------------------------------------------------------
## 1. RESHAPE: one row per (variable x condition) cell
## ---------------------------------------------------------------

conditions <- c("samt247", "MetA", "MetB", "MetC", "MetD")

# main "X vs gp120" direction columns (already include gp120 col separately)
main_cols <- c(
  gp120  = "GP120vsUNSTIM",
  samt247 = "GP120_samt247vsGP120",
  MetA   = "GP120_MetAvsGP120",
  MetB   = "GP120_MetBvsGP120",
  MetC   = "GP120_MetCvsGP120",
  MetD   = "GP120_MetDvsGP120"
)

# secondary "MetX vs samt247" columns (green over/under arrow), NA for gp120 & samt247
sub_cols <- c(
  gp120   = NA,
  samt247 = NA,
  MetA    = "MetAvssamt247",
  MetB    = "MetBvssamt247",
  MetC    = "MetCvssamt247",
  MetD    = "MetDvssamt247"
)

col_order <- c("gp120", "samt247", "MetA", "MetB", "MetC", "MetD")

long_main <- df %>%
  dplyr::select(variable, var_only, marker, all_of(main_cols)) %>%
  tidyr::pivot_longer(cols = all_of(col_order), names_to = "condition", values_to = "main_call")

long_sub <- df %>%
  dplyr::select(variable, all_of(na.omit(sub_cols))) %>%
  tidyr::pivot_longer(cols = -variable, names_to = "condition", values_to = "sub_call")


## ---------------------------------------------------------------
## 2. CLASSIFY EACH CALL INTO SYMBOL + COLOR
## ---------------------------------------------------------------

classify_main <- function(x) {
  case_when(
    is.na(x)                 ~ "none",
    grepl("^not_sig", x)     ~ "tilde",
    grepl("^up",     x)      ~ "up",
    grepl("^down",   x)      ~ "down",
    TRUE ~ "tilde"
  )
}

classify_sub <- function(x) {
  case_when(
    is.na(x)              ~ "none",
   # grepl("^not_sig", x)  ~ "none",   # no green arrow drawn when not significant
    grepl("^up",   x)     ~ "up",
    grepl("^down", x)     ~ "down",
    TRUE ~ "none"
  )
}

# background color per marker category (matches figure's pastel bands)
marker_colors <- c(
  "total" = "white",
  "IL17"  = "#E8D9F0",
  "CD107" = "#D6E4F5",
  "CD73"  = "#D3F0EC",
  "IL10"  = "#E6F2CC",
  "IFNg"  = "#FBE6C4",
  "TNFa"  = "#FBD6DA"
)


plot_df <- long_main %>%
  dplyr::left_join(long_sub, by = c("variable", "condition")) %>%
  dplyr::mutate(condition = factor(condition, levels = col_order),
                main_symbol = classify_main(main_call),
                sub_symbol  = classify_sub(sub_call),
                variable = factor(variable, levels = rev(df$variable)), # preserve original dataframe order
                bg_color = marker_colors[marker]
  )


## ---------------------------------------------------------------
## 3. BUILD THE PLOT
## ---------------------------------------------------------------

# tile background (colored by marker category)
p <- ggplot(plot_df, aes(x = condition, y = variable)) +
  geom_tile(aes(fill = bg_color), color = "grey40", linewidth = 0.3) +
  scale_fill_identity()

# main symbol: big arrow (red/black) or tilde (grey), size 6
main_shapes <- c(up = 24, down = 25, tilde = NA, none = NA)  # 24=filled up tri, 25=filled down tri

p <- p +
  # tilde symbol (text) for not_sig
  geom_text(
    data = filter(plot_df, main_symbol == "tilde", condition != "gp120"),
    aes(label = "~"), size = 7, color = "grey40", fontface = "bold"
  ) +
  # up arrow main (black if gp120 col, red otherwise per your legend: red = X+gp120 vs gp120; but
  # gp120 vs unstim uses open arrow style - handle gp120 column separately below)
  geom_text(
    data = filter(plot_df, main_symbol == "up", condition != "gp120"),
    aes(label = "\u2191"), size = 6, color = "red", fontface = "bold",
    nudge_y = 0.12
  ) +
  geom_text(
    data = filter(plot_df, main_symbol == "down", condition != "gp120"),
    aes(label = "\u2193"), size = 6, color = "black", fontface = "bold",
    nudge_y = 0.12
  ) +
  # gp120 column: open arrow style (up = gp120>unstim, tilde = gp120~unstim)
  geom_text(
    data = filter(plot_df, main_symbol == "tilde", condition == "gp120"),
    aes(label = "~"), size = 7, color = "grey60", fontface = "bold"
  ) +
  geom_text(
    data = filter(plot_df, main_symbol == "up", condition == "gp120"),
    aes(label = "\u2191"), size = 6, color = "grey60", fontface = "bold",
    nudge_y = 0.12
  ) +
  geom_text(
    data = filter(plot_df, main_symbol == "down", condition == "gp120"),
    aes(label = "\u2193"), size = 6, color = "grey60", fontface = "bold",
    nudge_y = 0.12
  )

# secondary green arrow overlay (Met-X vs SAMT-247), offset to upper-right of cell
p <- p +
  geom_text(
    data = filter(plot_df, sub_symbol == "up"),
    aes(label = "\u2191"), size = 5, color = "darkgreen", fontface = "bold",
    nudge_x = 0.28, nudge_y = 0.12
  ) +
  geom_text(
    data = filter(plot_df, sub_symbol == "down"),
    aes(label = "\u2193"), size = 5, color = "green3", fontface = "bold",
    nudge_x = 0.28, nudge_y = 0.12
  )

## ---------------------------------------------------------------
## 4. THEME / LABELS
## ---------------------------------------------------------------

p <- p +
  scale_x_discrete(position = "top",
                    labels = c("gp120", "SAMT-247", "MetA", "MetB", "MetC", "MetD")) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(face = "bold", size = 13),
    axis.text.y = element_text(size = 9),
    axis.ticks = element_blank(),
    legend.position = "none",
    plot.margin = margin(10, 10, 10, 10)
  )

plot(p)
  cat("\n")
        cat(' \n \n') ### this is the key!
  
```

```{r suppfig5print, results='asis', eval=TRUE}

  to_print <- sfig5_df %>%
    DT::datatable(extensions = 'Buttons',
                              options = list(dom = 'Blfrtip',
                                             buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                             iDisplayLength = 50,
                                             lengthMenu = list(c(10,25,50,-1),
                                                               c(10,25,50,"All"))))
              
              
              print(htmltools::tagList(to_print))


       cat("\n")
        cat(' \n \n') ### this is the key!
  


```












## Data Tables {.tabset}

### Raw Data

```{r rawdata_print}

#### add with button
xfun::embed_file(file.path(input_dir,paste0(str_before_last(params$study_name,"_"),"_list_of_RAWdatasets.xlsx")), text = "Excel spreadsheet of raw datasets")

```


### Plotted Values

```{r plottedValues_print}
##save
saveRDS(list_of_plotted, file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_plotted.rds")))
openxlsx::write.xlsx(list_of_plotted,file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_plotted.xlsx")))

#### add with button
xfun::embed_file(file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_plotted.xlsx")), text = "Excel spreadsheet of values plotted in each panel")

```

### Stats 

```{r stats_print}
##save
saveRDS(list_of_stats, file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_stats.rds")))
openxlsx::write.xlsx(list_of_stats,file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_stats.xlsx")))

#### add with button
xfun::embed_file(file.path(outputTABLES_dir,paste0(params$study_name,"_list_of_stats.xlsx")), text = "Excel spreadsheet of statistics for each panel")

```


## Session info

```{r sessioninfo}
sessionInfo()
```


